\LARGE \bf Compressed Directional Distance Transform for Fast 2D Ray Casting

### 2017-02-27 paper/main.tex
made large progress on initial editing pass

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphicx}
\usepackage{tabu}
\usepackage{float}
% \usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

@@

%: Margin note commands
 \usepackage[usenames,dvipsnames]{xcolor}
\setlength{\marginparwidth}{0.5in}
\let\oldmarginpar\marginpar
\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\footnotesize #1]%
{\raggedright\footnotesize #1}}
\newcommand{\skcomment}[2]{{\color{red}#1} \marginpar{\color{TealBlue}\tiny #2 (SK)}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

\skcomment{}{Perhaps start with a short introduction to localization, and then describe what particle filter (Monte Carlo) localization is, and then introduce ray casting in that context.}

The ray cast operation is an important sub-problem for robot localization \cite{localization} using distance sensors to detect the environment. In particle filter localization, particles are re-weighted according to a Bayesian update equation which relies on ray casting to determine the ground truth distance between the hypothesis pose and obstacles in a map. Effective particle filters often maintain thousands of particles, each of which must be reweighted tens of times per second, and each re-weighting may require tens of ray casts. As a result, millions of ray cast operations may be resolved per second, posing a significant computational challenge for resources constrained systems.

To combat this challenge, \skcomment{Fox et al.}{Thrun et al.?} \cite{localization} suggest the use of a large three-dimensional lookup table (LUT) to store expected ranges for each discrete $(x,y,\theta)$ state. While this is simple to implement and does result in large speed improvements as compared to ray casting, it can be prohibitively memory intensive for large maps and/or resource constrained systems\skcomment{.}{perhaps give an example: In such and such large area with such and such resolution map, the data structure is XX GigaBytes (TeraBytes?)}

We propose an analogous data structure called a Compressed Directional Distance Transform (CDDT) which requires significantly less memory and precomputation time while still allowing near constant time ray casting queries.

\skcomment{}{This is probably a good place to list the contributions of this work as in: ``The contributions of this work are the following. Firstly, we develop ... Secondly, we implement this algorithm for a ... You can also say things like: We observe two orders of magnitude improvement in memory requirements, with little sacrifice in computation time, when compared to lookup table methods. Et cetera.}

\skcomment{Many robotic applications require three dimensional maps, however that is considered outside the scope of this work since many data structures exist for accelerating 3D ray casting, such as the bounded volume hierarchy and related variants.}{I would put the discussion in this paragraph to the future work section in the conclusions, rather than here.} The five dimensional table required to precompute expected ranges for a discrete state space $(x,y,z,\theta,\phi)$ would be prohibitively expensive to both compute and store, so 3D acceleration data structures are exclusively used when 3D ray casting is required. \skcomment{Though our approach does extend to higher dimensionality and would be smaller than the analogous five dimensional table, it would still be prohibitively expensive in greater than two dimensions, thus we do not consider it here.}{I would not worry about putting this justification either.}

\skcomment{
\section{RELATED WORK}
}{You might want to integrate this section into the introduction section. Otherwise the introduciton section is unusually short. Many people discuss related work in the introduciton. You can put this discussion right after your second paragraph, and right before listing all your contributions.}
Several widely used algorithms exist for ray casting in a two dimensional space. Due to varied handling of edge cases and ambiguities in the problem of determining distance between any point and the nearest obstacle in a particular direction, most 2D ray casting algorithms do not provide exactly consistent results for every query. However, on average error between results from each of the methods discussed is small.

\skcomment{\subsection{Bresenham's Line}}{I would not worry about these section headings. You say the names of the algorithms in the beginning of the paragraphs anyways.}

Bresenham's line algorithm \cite{bresenham} incrementally determines the set of pixels that approximate the trajectory of query ray starting from the query point $(x,y)$ and progressing in the $\theta$ direction. The algorithm terminates once the nearest occupied pixel is discovered, and the euclidean distance between that occupied pixel and the query $(x,y)$ is reported. This algorithm is widely implemented in particle filters due to its simplicity and ability to operate on a dynamic map. The primary disadvantage is that it is slow, potentially requiring hundreds of memory accesses for a single ray cast. While actual performance is highly environment dependent, Bresenham’s Line algorithm is linear in map size in the worst case.

\subsection{Ray Marching Distance Transforms}

Similar to Bresenham's Line algorithm, ray marching \cite{raymarching} checks points along the line radiating outwards from the query point until an obstacle is encountered. The primary difference is that ray marching makes larger steps along the query ray, avoiding unnecessary memory reads. Beginning at the query point $(x,y)$ the ray marching algorithm proceeds in the theta direction, stepping along the line by the minimum distance between each query point and the nearest obstacle. The algorithm terminates when the query point falls on an obstacle in the map. A precomputed euclidean distance transform of the map provides the distance between each query point and the nearest obstacle in any direction.

@@

Ray marching is on average much faster than Bresenham’s line, but edge cases exist in which the performance is equivalent. For example, a query ray which travels closely parallel to a wall will make small steps between each query point for the whole length of the ray. As a result, the performance of ray marching \skcomment{has a long tail distribution}{is this widely known? Is there a paper that you can cite here? Or can you refer the reader to your results in this paper (e.g., write ``(See Section X.)'' to tell the reader that there are some results there on heavy tails of this algorithm.} which can be problematic for near real time algorithms. Though highly scene dependent, on average ray marching runtime is roughly logarithmic in distance between obstacles, though in the worst case it is linear in map size.

% \begin{figure}[thpb]
%       \centering
%       \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
% }}
%       %\includegraphics[scale=1.0]{figurefile}
%       \caption{Inductance of oscillation winding on amorphous
%       magnetic core versus DC bias magnetic field}
%       \label{figurelabel}
%   \end{figure}

\subsection{Lookup Table}

As previously described, a common acceleration technique for two dimensional ray casting is to simply precompute the ray distances for every state in a discrete grid and store the results in a three dimensional LUT. Any ray casting technique may be used to compute the table. In the authors’ implementation ray casting is used to populate the LUT since it has low initialization time and is significantly faster than Bresenham’s Line. Theoretically this approach has constant query runtime, though actual performance is dependent on access patterns as CPU caching effects are significant in practice.

State space discretization implies approximate results, since intermediate states must be rounded onto the discrete grid. While the effect is small for x and y, it can be significant for theta. As the ray moves further away from the query point along the theta direction, angular discretization error accumulates. For queries $(x,y,\theta)$ discretized into $(x_d,y_d,\theta_d)$, the distance between the end of the ray $(x_d,y_d,Θ_d)$ and its projection onto the line implied by $(x,y,\theta)$ becomes large as the length of the ray increases. Although the error induced by discretization may be unacceptable for computer graphics applications, it is generally acceptable for probabilistic algorithms such as the recursive Bayes filter since error is expected and directly modeled.

One method to improve accuracy of queries which fall between discrete states is to ray cast the neighboring discrete states and interpolate between them, rather than simply rounding to the nearest discrete state. Since we are evaluating the performance of these methods for use in a particle filter, we do not perform interpolation since the increased computation reduces the number of particles that may be maintained in real time and therefore any increase in accuracy with respect to the stored map is effectively negated.

\skcomment{
\section{APPROACH}
}{Instead of this section title, i would use ``The Compressed Directional Distance Transform (CDDT) Algorithm'' or something like that.}

Although the three dimensional table used to store precomputed ray cast solutions for a discrete state space is inherently large, it is highly compressible. This is most apparent in the cardinal directions, in which adjacent values along a particular dimension of the table increase by exactly one unit of distance for unobstructed positions. Our data structure is designed to compress this redundancy while still allowing for fast queries in near constant time. We accomplish this though through what we refer to as a compressed directional distance transform (CDDT) described here.

\begin{figure}[h]
\img{basement_bw_cropped.png}
\caption{Occupancy grid map of the Stata basement loop. 1200x1200}
\label{stata}
\end{figure}

The euclidean distance transform (EDT) of an occupancy grid map stores the distance to the nearest obstacle in any direction for each point $(x,y)$ in the map. In contrast, a directional distance transform (DDT) stores the distance to the nearest obstacle in a particular direction for each point $(x,y)$ in the map. The key difference between a single Θ slice of the LUT and a directional distance transform is the way in which it is computed. To compute the LUT slice, ray casting is performed in the $\theta_d$ direction from every $(x_d,y_d)$ in the map using another algorithm such as Bresenham’s line. In contrast, to compute the DDT, the obstacles in the map are rotated about the center of the map by $-\theta_d$ and the values of each column in the DDT may be directly computed as a sawtooth function which goes to zero whenever an obstacle is encountered in that column. \skcomment{}{Do you think this is a good place for an example?} Thus, each column in the DDT may be characterized as a sawtooth function with \skcomment{uneven periodicity}{what is uneven periodicity mean?}. While it is true that the rotation of the occupancy grid introduces small errors as compared to ray casting for each cell, as previously discussed small errors do not significantly impact particle filter localization performance due to the use of a probabilistic error model. In practice, errors induced by map rotation are dominated by those inherently present in distance sensors.

\begin{figure}[h]
\img{slice0_cddt_basement.png}
\caption{A single slice of the DDT generated from the CDDT data structure for $\theta=0$. The value of each pixel represents the distance to the nearest obstacle in the upwards direction with respect to the image above. Each column may be thought of as a sawtooth wave, and the points where the wave is zero are projected into the CDDT bin for that column for later indexing. The blue plot indicates the number of zero points projected into each CDDT bin. The compression factor of 267 is computed as two times the size of the corresponding LUT slice divided by the size of this slice in the CDDT. The factor of two is due to the use of rotational symmetry to reduce the number of slices maintained in the CDDT with respect to the LUT.}
\label{cddt:slice0}
\end{figure}

\begin{figure}[h]
\img{slice41_cddt_basement.png}
\caption{Another slice of the DDT generated from the CDDT data structure for $\theta=2.385$.}
\label{cddt:slice41}
\end{figure}

\skcomment{The conversion from a DDT to a CDDT is performed by converting the sawtooth response of each column into a set of points in which the sawtooth wave goes to zero.}{This is probably the most important part, but kind of hard to understand. I would consider an example here as well.} This conversion is lossless in the sense that each element of the table may be recovered as a function of the nearest zero point in a particular direction. If the set of zero points is sorted, then the magnitude of the saw tooth wave at any point in the original DDT may be quickly discovered by performing binary search for the nearest zero point in that column and finding the distance between the query point and that zero point. This losslessness and fast indexing allows for ray cast queries to be quickly resolved without storing the entire DDT, but rather by storing the sorted lists of zero points.
For sake of efficiency, the full DDT need not be computed, but rather the CDDT may be built directly by projecting each obstacle into the coordinate space of each theta slice and storing a list of y-coordinates for each discrete x-coordinate. Once all scene geometry is projected into the table, each list of zero points is sorted to facilitate indexing.

\skcomment{}{This section could be organized a bit better. Perhaps have a subsection that says ``Compression algorithm" (text above here) and another one that says ``Query Algorithm'' (text below here).}

@@

\subsection{Rotational Symmetry}

@@

\begin{figure}[h]
\img{particle_trial2_basement_cropped.png}
\caption{Violin plots for ray casting completion times during the execution of a particle filter. Basement map. Ray casting is primarily done inside of hallways, reducing the mean ray length and thereby improving Bresenham's Line performance. CDDT and PCDDT exploiting radial symmetry while resolving the sensor model.}
\label{violin:basement:random}
\end{figure}

@@

>>>>

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphicx}
\usepackage{tabu}
\usepackage{float}
% \usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{tikz}
\usetikzlibrary{matrix}
\usetikzlibrary{shapes.geometric, arrows}

@@

\newcommand{\gridScale}{1.25}

%: Margin note commands
\usepackage[usenames,dvipsnames]{xcolor}
\setlength{\marginparwidth}{0.5in}
\let\oldmarginpar\marginpar
\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\footnotesize #1]%
{\raggedright\footnotesize #1}}
\newcommand{\skcomment}[2]{{\color{red}#1} \marginpar{\color{blue}\tiny #2 (SK)}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Determining a robot's location and orientation in a known environment, also known as localization, is an important and challenging problem in the field of robotics. Particle filters are a popular class of Monte Carlo algorithms used to track the pose of mobile robots by iteratively refining a set of pose hypotheses called particles. After determining an initial set of particles, the particle filter updates the position and orientation of each particle by applying a movement model based on available odometry data. Next, the belief in each particle is updated by comparing sensor readings to a map of the environment. Finally, the particles are resampled according to the belief distribution and algorithm repeats.

While particle filters provide a robust and general framework for solving the localization problem, they can be computationally expensive due to both the number of particles which must be maintained and the evaluation of the sensor model. In robots with range sensors such as LiDAR or Sonar, ray casting is frequently used to compare sensor readings with the ground truth distance between the hypothesis pose and obstacles in a map. Ray casting itself is a complex operation, and reweighting any single particle may require tens of ray casts. Many effective particle filters maintain thousands of particles, updating each particle tens of times per second. As a result, millions of ray cast operations may be resolved per second, posing a significant computational challenge for resource constrained systems.

Several well known algorithms exist for ray casting in two dimensional spaces such as Bresenham's Line algorithm \cite{bresenham} and ray marching \cite{raymarching}. Both algorithms work by iteratively checking points in the map starting at the query point and moving in the ray direction until an obstacle is discovered. This process may require hundreds of memory reads per ray cast depending on the distance to the nearest obstacle, and does not provide constant time performance.

To combat the computational challenges of ray casting while localizing in a two-dimensional map, Thrun et al. \cite{localization} suggest the use of a large three-dimensional lookup table (LUT) to store expected ranges for each discrete $(x,y,\theta)$ state. While this is simple to implement and does result in large speed improvements as compared to ray casting, it can be prohibitively memory intensive for large maps and/or resource constrained systems. In a 2000 by 2000 occupancy map, storing ranges for 200 discrete directions would require over 1.5GB. While this memory requirement may be acceptable in many cases, it scales with the square of map size - a 4000 by 4000 map would require over 6GB for the same angular discretization.

We propose an analogous data structure to the three-dimensional LUT called a Compressed Directional Distance Transform (CDDT) which requires significantly less memory and precomputation time while still allowing near constant time ray casting queries. The contributions of this work are the following. Firstly, we develop a problem formulation and provide details of our method. Secondly, we implement the proposed algorithm and compare its performance to alternate two-dimensional ray casting methods. Finally, we discuss potential extensions to the CDDT algorithm which could be the basis of future research. We observe two orders of magnitude improvement in memory requirements, with little sacrifice in computation time, when compared to the lookup table methods. Additionally, we observe a large speedup when compared to the other ray casting methods considered with similar memory requirements.

\section{Related Work}

Due to varied handling of edge cases and ambiguities in the problem of determining distance between any point and the nearest obstacle in a particular direction, most 2D ray casting algorithms do not provide exactly consistent results for every query. However, on average error between results from each of the methods discussed is small.

Bresenham's line algorithm \cite{bresenham} incrementally determines the set of pixels that approximate the trajectory of query ray starting from the query point $(x,y)$ and progressing in the $\theta$ direction one pixel at a time. The algorithm terminates once the nearest occupied pixel is discovered, and the euclidean distance between that occupied pixel and the query $(x,y)$ is reported. This algorithm is widely implemented in particle filters due to its simplicity and ability to operate on a dynamic map. The primary disadvantage is that it is slow, potentially requiring hundreds of memory accesses for a single ray cast. While actual performance is highly environment dependent, Bresenham’s Line algorithm is linear in map size in the worst case.

Similar to Bresenham's Line algorithm, ray marching \cite{raymarching} checks points along the line radiating outwards from the query point until an obstacle is encountered. The primary difference is that ray marching makes larger steps along the query ray, thereby avoiding unnecessary memory reads. Beginning at the query point $(x,y)$ the ray marching algorithm proceeds in the theta direction, stepping along the line by the minimum distance between each query point and the nearest obstacle. The algorithm terminates when the query point coincides with an obstacle in the map. A precomputed euclidean distance transform of the occupancy map provides the distance between each query point and the nearest obstacle in any direction.

@@

Ray marching is on average much faster than Bresenham’s line, but edge cases exist in which the performance is nearly equivalent. As noted in \cite{acceleratedraymarching} the traversal speed of rays rapidly decreases as sampled positions approach obstacles. For this reason, rays which travel parallel to walls progress very slowly as compared to those passing through open areas. Thus the performance of ray marching exhibits a long tail distribution as seen in (Fig. \ref{violin:basement:all}, \ref{violin:basement:particle}) which negatively impacts average runtime as compared to the median, and can be problematic for near real time algorithms.

As previously described, a common acceleration technique for two dimensional ray casting is to precompute the ray distances for every state in a discrete grid and store the results in a three dimensional LUT. Any ray casting technique may be used to compute the table. In the authors’ implementation ray casting is used to populate the LUT since it has low initialization time and is significantly faster than Bresenham’s Line. Theoretically this approach has constant query runtime, though actual performance is dependent on access patterns as CPU caching effects are significant in practice.

State space discretization implies approximate results, since intermediate states must be rounded onto the discrete grid. While the effect of rounding query position, it can be significant for $\theta$. As the ray moves further away from the query point along the theta direction, angular discretization error accumulates. For queries $(x,y,\theta)$ discretized into $(x_d,y_d,\theta_d)$, the distance between the end of the ray $(x_d,y_d,Θ_d)$ and its projection onto the line implied by $(x,y,\theta)$ becomes large as the length of the ray increases. Although the error induced by discretization may be unacceptable for computer graphics applications, it is generally acceptable for probabilistic algorithms such as the particle filter since error is expected and directly modeled.

Rather than simply rounding to the nearest grid state, one method to improve accuracy of queries which lie between discrete states is to query the neighboring states and interpolate between them. Since we are evaluating the performance of these methods for use in a particle filter, we do not perform interpolation since the increased computation reduces the number of particles that may be maintained in real time, and therefore any increase in accuracy with respect to the stored map is effectively negated.

\section{Problem Formulation and Notation}

We define the problem of ray casting in occupancy grids as follows. We assume a known occupancy grid map in which occupied cells have value one and unoccupied cells have value zero. Given a query pose $(x,y,\theta)_{query}$ in map space, the ray cast operation finds the nearest occupied pixel $(x,y)_{collide}$ which lies on the ray starting at the position $(x,y)_{query}$ pointing in the $\theta_{query}$ direction. The ray cast operation returns the euclidean distance $d_{ray}$ between $(x,y)_{query}$ and $(x,y)_{collide}$.

$$ d_{ray} = \left\Vert\begin{pmatrix}
           x \\
           y 
         \end{pmatrix}_{query} - 
         \begin{pmatrix}
           x \\
           y
         \end{pmatrix}_{collide}\right\Vert_2
$$

We refer to $(x,y,\theta)_{query}$ as the query pose, and $(x,y)_{query}$ as the query point. Additionally, denote the discretized query pose as $\lfloor(x,y,\theta)_{query}\rceil$. A $\theta$ slice through the LUT refers to all table entries for that particular $\lfloor\theta\rceil$.

\begin{figure}[h]
\img{coord_system.png}
\caption{Occupancy grid map coordinate system}
\label{coordinate_system}
\end{figure}

\section{The Compressed Directional Distance Transform Algorithm}

Although the three dimensional table used to store precomputed ray cast solutions for a discrete state space is inherently large, it is highly compressible. This is most apparent in the cardinal directions, in which adjacent values along a particular dimension of the table increase by exactly one unit of distance for unobstructed positions as in Fig. \ref{cardinal_example}. Our data structure is designed to compress this redundancy while still allowing for fast queries in near constant time. We accomplish this though through what we refer to as a Compressed Directional Distance Transform (CDDT) described here.

\begin{figure}[h]
\begin{center}
\begin{tikzpicture}
\fill[black,scale=\gridScale] (0.5,0.5) rectangle (1,1);
\fill[black,scale=\gridScale] (-0.5,0) rectangle (0,0.5);
\fill[black,scale=\gridScale] (0,-0.5) rectangle (0.5,0);
\fill[black,scale=\gridScale] (0.5,-1) rectangle (1,-0.5);
\draw[scale=\gridScale,step=0.5cm,color=white] (-1,-1.1) grid (1,1.1);
\draw[scale=\gridScale,step=0.5cm,color=gray] (-1,-1) grid (1,1);
\end{tikzpicture}
\begin{tikzpicture}
\draw[scale=\gridScale,step=0.5cm,color=white] (-1,-1.1) grid (1,1.1);
\draw[scale=\gridScale,step=0.5cm,color=gray] (-1,-1) grid (1,1);
\matrix[matrix of nodes,nodes={inner sep=0pt,text width=.5cm,align=center,minimum height=0.45cm, scale=\gridScale}]{
3 & 2 & 1 & 0 \\
1 & 0 & - & - \\
2 & 1 & 0 & - \\    
3 & 2 & 1 & 0\\};
\end{tikzpicture}
\end{center}
\caption{Occupancy grid (left) and associated LUT for $\theta=0$ (right).}
\label{cardinal_example}
\end{figure}

The euclidean distance transform of an occupancy grid map stores the distance to the nearest obstacle in any direction $\theta$ for each possible $\lfloor(x,y)\rceil$ in the map. In contrast, a directional distance transform (DDT) stores the distance to the nearest obstacle in a particular direction $\lfloor\theta\rceil$ for every $\lfloor(x,y)\rceil$. The key difference between a single $\theta$ slice of the LUT and a directional distance transform for the same $\theta$ is the way in which it is computed and indexed. To compute the LUT slice, ray casting is performed in the $\theta$ direction for every $\lfloor(x,y)\rceil$. At runtime, each $(x,y)_{query}$ is discretized to $\lfloor(x,y)_{query}\rceil$ and the LUT slice is directly indexed.

In contrast, to compute the DDT, the obstacles in the map are rotated about the origin by $-\lfloor\theta\rceil$ and ray casting is implicitly performed in the $\theta=0$ direction to populate the table. Then, to index the table for a query $(x,y,\theta)_{query}$ one discretizes to $\lfloor(x,y,\theta)_{query}\rceil$, rotates $\lfloor(x,y)_{query}\rceil$ about the origin, and uses the rotated coordinates $\lfloor(x,y,\theta)_{query}\rceil_{rot}$ to index into the DDT. Thus, the same operation is computed in both the DDT and the LUT, but while one changes the ray cast direction to populate the LUT, one rotates scene geometry and ray casts in a constant direction to populate the DDT. While it is true that the rotation of the scene geometry introduces small errors as compared to ray casting for each cell, as previously discussed small errors do not significantly impact particle filter localization performance due to the use of a probabilistic error model and are dominated by errors inherently present in distance sensors.

The distinction between the LUT slice and the DDT may be subtle, but it has an important effect. Since ray casting is always performed in the $\theta=0$ direction to populate the DDT, all values in the same row of the DDT either increase by one unit with respect to their neighbor in the $\theta=0$ direction or go to zero. Thus each row of the DDT may be characterized as a sawtooth function where the zero points correspond to obstacles in the map. This characterization as a sawtooth function provides a natural method of lossless compression: keep the zero points and discard the rest.

The conversion from a DDT to a CDDT slice is performed by storing the $x$ coordinate of each zero point in the coordinate space of the DDT for every discrete row. At query time, the sawtooth function (which is the distance to the nearest obstacle in the query direction) encoded in the DDT may be quickly recovered by performing binary search for nearest zero point in the correct row of the CDDT slice. We refer to the list of zero points for a single row as a CDDT bin. Similar to the LUT, the full CDDT is defined as every CDDT slice for all discrete values of $\theta$.

For performance, it is not necessary to compute the full DDT, but rather the CDDT may be directly constructed by projecting each obstacle into the coordinate space of the DDT for every $\lfloor\theta\rceil$ and sorting its $x$ coordinate in the CDDT bin corresponding to its $y$ coordinate and rotation. After all geometry has been projected into the CDDT, each bin is sorted to facilitate later indexing. Not only does the direct construction of the CDDT greatly reduce the amount of memory required to store the LUT, it also reduces data structure precomputation time since ray casting is not necessary.

[TODO add an example of scene rotation, saw tooth wave, and zero point projection]

\subsection{CDDT Construction Algorithm}

\subsection{CDDT Query Algorithm}

{\color{red}DONE WITH SECOND EDITING PASS UP TO THIS POINT}

@@

\subsection{Further Optimizations}

@@

\begin{figure}[h]
\img{particle_trial2_basement_cropped.png}
\caption{Violin plots for ray casting completion times during the execution of a particle filter. Basement map. Ray casting is primarily done inside of hallways, reducing the mean ray length and thereby improving Bresenham's Line performance. CDDT and PCDDT exploiting radial symmetry while resolving the sensor model.}
\label{violin:basement:particle}
\end{figure}

@@

\bibitem{acceleratedraymarching} Zuiderveld, Karel, Koning, Anton, and Viergever, Max, “Acceleration of ray-casting using 3D distance transforms”, Proceedings of the SPIE – Visualization in Biomedical Computing 1992, Vol.1808, pp.324-335, 1992.
